Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-03-19 index.qmd
minor fix to index.qmd

---
title: 'Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
# subtitle: 'The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
shorttitle: 'Effect of Recruiting Military Veterans on Confidence in 2024 US Elections' # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
date: "2024-10-16"
date-modified: today
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

# load packages
library(tidyverse)
library(modelsummary)
library(tinytable) # simple tables in html, LaTeX, markdown, word, pdf, typst
library(litedown)
# library(MASS)
# library(ordinal)
# library(marginaleffects)
# library(easystats)
# library(patchwork)

@@

# apply Quarto data processing to all tables
# theme_quarto <- function(x) format_tt(x, quarto = TRUE)
# options(tinytable_tt_theme = theme_quarto)

@@

# load data dictionary/codebook
codebook <- read.csv("data/survey_codebook.csv")
data_dict <- read.csv("data/data_dictionary.csv")

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned[^2]. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate.

@@

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Later, I compare and discuss results between the items pertaining to different settings.

@@

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's favorability to local election officials ("In general, how favorable or unfavorable is your impression of local election officials?") and legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). The former presented five response options from "Strongly favorable" to "Strongly unfavorable"; the latter item presented a binary of either "Legitimate" or "Not Legitimate".

@@

Finally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^8].

@@

# kableExtra::kbl(conf_items_text, booktabs = T,
#                 col.names = c("Trust", "Distrust"),
#                 linesep = "\\addlinespace") |> 
#   kableExtra::kable_styling(full_width = T,
#                             font_size = 8,
#                             latex_options = c("HOLD_position")) |> 
#   kableExtra::column_spec(column = 1:2, width = "8cm")

@@

# output table as `kable` for LaTeX pdf
# kableExtra::kbl(
#   tab_ttest,
#   booktabs = T,
#   col.names = c(
#     "Place",
#     "$\\bar{x}_{diff}$",
#     "$\\bar{x}_{treat}$",
#     "$\\bar{x}_{control}$",
#     "$treat_{n}$",
#     "$control_{n}$",
#     "t",
#     "p",
#     "data",
#     "CI"
#   ),
#   escape = F
# ) |>
#   kableExtra::footnote(general = "Two Sample t-tests of difference between Confidence in Elections by Treatment Condition.") |>
#   kableExtra::kable_styling(font_size = 9, protect_latex = T)

@@

  # kableExtra::kbl(tab_q25_pdiff, 
  #                 booktabs = T) |> 
  # kableExtra::footnote(general = "Results derived from proportional odds logistic regression model. Estimates are differences in predicted probabilities transformed from log odds ratios, and represent difference in estimated prediction of choosing a particular response or higher")
```

@@

  # kableExtra::kbl(tab_q26_pdiff, 
  #                 booktabs = T) |> 
  # kableExtra::footnote(general = "Results derived from proportional odds logistic regression model. Estimates are differences in predicted probabitilites transformed from log odds ratios, and represent difference in estimated prediction of choosing a particular response or higher")

@@

# 
# Response <- data |> 
#   dplyr::select(q7, q25, q26, q36, q37) |>
#   surveytoolbox::data_dict() |> 
#   dplyr::as_tibble() |> 
#   dplyr::select(value)
# 
# 
# item_table <- data |> 
#   select(q7, q25, q26, q36, q37) |>
#   dplyr::rename_with(.cols = everything(), .fn = ~snakecase::to_any_case(.,case = "upper_camel")) |> 
#   var_label_tab() |>
#   select(Question = var_label) |> 
#   mutate(Response = Response$value)
# 
# item_table
# 
# tinytable::tt(item_table, width = c(.7, .3)) |> 
#   tinytable::style_tt(fontsize = 0.80)

# alternate
# kableExtra::kbl(item_table, booktabs = T, linesep = "\\addlinespace[2mm]") |> 
#   kableExtra::kable_styling(
#     full_width = T,
#     font_size = 9,
#     latex_options = c("HOLD_position")
#     ) |> 
#   kableExtra::pack_rows(index = c("Legitimacy" = 1,
#                                   "Safety Concerns" = 4))

@@

[^8]: Survey item wording and responses are included in Appendix A.

>>>>

---
title: 'Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
# subtitle: 'The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
shorttitle: 'Effect of Recruiting Military Veterans on Confidence in 2024 US Elections' # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
abstract: |
  In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

# load packages
library(tidyverse)
library(modelsummary)
library(tinytable) # simple tables in html, LaTeX, markdown, word, pdf, typst
library(ordinal)
library(marginaleffects)
library(easystats)
library(patchwork)

@@

@@

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned[^2]. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.

@@

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. I compare and discuss results between the items pertaining to different settings.

@@

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*. Additionally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^8]. Although two survey items are insufficient to serve as a composite scale, safety concerns will be evaluated by examination of response to these two questions in particular.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's beliefs about the legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). Respondents selected either "Legitimate" or "Not Legitimate".

@@

@@

@@

@@

@@

@@

@@

[^8]: Question wording and responses to a selection of survey items are included in Appendix A.
